---
title: "class14"
author: "Rose Li"
date: "5/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import our countData and colData into R

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

Q1. How many genes are we dealing with here 

```{r}
nrow(counts)
```

Lets check if the cols of 'counts' is the same as the id column in 'metadata'
```{r}
colnames(counts)
```

```{r}
metadata$id
```
```{r}
colnames(counts)[-1]
```

```{r}
colnames(counts)[-1] == metadata$id
```

```{r}
all(colnames(counts)[-1] == metadata$id)
```

## Differential Expression 

Lets access just their control counts data so we can calculate an average counts value for each gene in these control samples 
```{r}
control <- metadata[metadata$dex=="control",]
control
```
Now we want to get the cols in counts with these 'control$id' 

```{r}
control.mean <- rowSums( counts[ ,control$id] )/length(control$id)
head(control.mean)
```


```{r}
names(control.mean) <- counts$ensgene
head(control.mean)
```

Now do the same thing for the treatment 

```{r}
treated <- metadata[metadata$dex=="treated",]
treated.mean <- rowSums( counts[ ,treated$id] )/length(treated$id)
names(treated.mean) <- counts$ensgene
head(treated.mean)
```

```{r}
meancounts <- data.frame(control.mean, treated.mean)
head(meancounts)
```

```{r}
plot(control.mean, treated.mean)
```


Plot on log scale for both axis
```{r}
plot(meancounts, log="xy")
```

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

```{r}
# identify the rows (i.e. genes) with zero values 
to.rm <- unique(which(meancounts[,1:2]==0, arr.ind=TRUE)[,1])

# remove these rows as we can't really say much about them 
mycounts <- meancounts[-to.rm,]
head(mycounts)
```



```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

```{r}
up.ind <- mycounts$log2fc > 2
sum(up.ind==TRUE)
down.ind <- mycounts$log2fc < (-2)
sum(down.ind==TRUE)
```

## Adding annotation data 
```{r}
anno <- read.csv("annotables_grch38.csv")
head(anno)
```

We will use the **merge()** function to add annotation data to our 'mycounts' results 

```{r}
merge(x=mycounts, y=anno, by.x="row.names", by.y="ensgene")
```

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

```{r}
columns(org.Hs.eg.db)
```

```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
```

```{r}
mycounts$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")
```
```{r}
mycounts$uniprot < mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="UNIPROT",
                     keytype="ENSEMBL",
                     multiVals="first")
```

## DESeq2 Analysis 
```{r}
library(DESeq2)
citation("DESeq2")
```

Build our deseq dataset object needed for running deseq analysis 
```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```


```{r}
dds <- DESeq(dds)
```

Get our results 
```{r}
res <- results(dds)
as.data.frame(res)
```

```{r}
summary(res)
```

```{r}
res05 <- results(dds, alpha = 0.05)
summary(res05)
```

```{r}
resSig05 <- subset(as.data.frame(res), padj < 0.05)
nrow(resSig05)
```


```{r}
resSig01 <- subset(as.data.frame(res), padj < 0.01)
nrow(resSig01)
```

```{r}
ord <- order( resSig01$padj )
#View(res01[ord,])
head(resSig01[ord,])
```

```{r}
write.csv(resSig01[ord,], "signif01_results.csv")
```


## Visualization 
```{r}
res$sig <- res$padj<0.05 & abs(res$log2FoldChange)>2
table(res$sig)
```

```{r}
sum(is.na(res$sig))
```

```{r}
# Set the color palette for our plot
palette( c("gray","blue") )

plot( res$log2FoldChange,  -log(res$padj), 
 col=res$sig+1, ylab="-Log(P-value)", xlab="Log2(FoldChange)")

# Add some cut-off lines
abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.1), col="darkgray", lty=2)
```


```{r}
# Setup our custom point color vector 
mycols <- rep("gray", nrow(res))
mycols[ abs(res$log2FoldChange) > 2 ]  <- "red" 

inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

#Volcano plot with custom colors 
plot( res$log2FoldChange,  -log(res$padj), 
 col=mycols, ylab="-Log(P-value)", xlab="Log2(FoldChange)" )

abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.1), col="gray", lty=2)
```


```{r}
library(ggplot2)
ggplot(as.data.frame(res), aes(log2FoldChange, -log10(pvalue), col=sig)) + 
    geom_point() + 
    ggtitle("Volcano plot")
```

Another way to plot this with the Enhanced 

But first lets put proper gene names on our results (i.e. gene 'symbols')

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
x <- as.data.frame(res)
x$symbols <- mapIds(org.Hs.eg.db,
                    keys=row.names(x),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
```


```{r}
library(EnhancedVolcano)
EnhancedVolcano(res,
    lab = x$symbols,
    x = 'log2FoldChange',
    y = 'pvalue')
```

